---
title: "Script linked to Figure 6:  Gradual blockade of pallial lineage neuronal differentiation within the postnatal dorsal SVZ."
subtitle: "Figure 6A, 6B, 6C, 6D, 6G, 6H, 6I, 6K"
output:
  rmdformats::downcute:
    self_contained: true
    default_style: "light"
    downcute_theme: "chaos"
---
***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load packages

```{r}
suppressWarnings(suppressMessages({
library(tidyverse,quietly=T)
library(Seurat,quietly=T)
library(patchwork,quietly=T)
library(viridis,quietly=T)
library(scales,quietly=T)
library(treemap,quietly=T)
library(ggplot2,quietly=T)
library(ggpubr,quietly = T)
library(Rmagic, quietly = T)
library(reticulate, quietly=T)
library(magrittr, quietly=T)
library(imager, quietly=T)
library(EBImage, quietly=T)
library(STutility, quietly=T)
library(magrittr, quietly=T)
library(dplyr, quietly=T)
library(DT, quietly=T)
library(kableExtra, quietly=T)
}))
```

```{r}
CLBNxs_Neurons_new_idents_V2=readRDS("C:/Users/raine/OneDrive/Documents/BioInfo/neuronalcells/CLBNxs_Neurons_new_idents_V2.RDS")
```

# A

>Fig6A: UMAP plot with identity of GLU (deep red gradient) and GABA (deep blue gradient) TAPs/NBs subclusters at P12.

```{r,fig.width=4,fig.height=3,fig.align="center"}
DimPlot(CLBNxs_Neurons_new_idents_V2,pt.size = 2, label = F, cols = c("#5A0017","#800021","#B76E78","#e4c1c6","#001277","#0c409f","#3074c7","#70a1dc"))
```

# B

>Fig6B: Integration of current P12 dataset with previously published dataset of postnatally born olfactory bulb neurons (GSE134918, gray).

```{r,fig.height=5,fig.width=6,fig.align="center"}

Mizrak2020_CLB=readRDS("C:/Users/raine/OneDrive/Documents/BioInfo/mizrak2020/Mizrak2020_CLB.rds")

Mizrak2020_CLB <- RenameIdents(Mizrak2020_CLB,  '5' = 'GLU1', '3' = 'GLU2', '7' = 'GLU3','4' = 'GABA1','0' = 'GABA2','2' = 'GABA3','1' = 'GABA4' )
p=DimPlot(Mizrak2020_CLB, pt.size = 2, cols = c("#5A0017","#800021","#B76E78","#001277","#0c409f","#3074c7","#70a1dc","#e4c1c6","#dedede","#dedede","#dedede","#dedede","#dedede"),label=F)
LabelClusters(p,id="ident",fontface="bold")+NoLegend()
```

# C

>Fig6C: Feature plot indicating cycle phase, S.Score and G2M.Score of GLU and GABA cells.

```{r,fig.width=4,fig.height=3,fig.align="center"}
DimPlot(CLBNxs_Neurons_new_idents_V2,pt.size = 2, label = F, group.by = "Phase",cols = c("#dad6cd","#a3b28a","#588157"))
```

# D

>Fig6D: Representation of overall gene count per feature as revealed by Visium spatial transcriptomics within the SVZ at corresponding timepoints.

```{r, fig.width=10, fig.height=3,fig.align="center"}

se.cropped = readRDS("C:/Users/raine/OneDrive/Documents/visium/merge_se_cropped.rds")
se.cropped <- suppressMessages(suppressMessages(se.cropped %>% 
  SCTransform(verbose = F) %>%
  RunPCA(verbose=F) %>%
  RunUMAP(reduction = "pca", dims = 1:20,verbose = F)))

FeatureOverlay(se.cropped, features = "Dlx2",
sampleids = 2:4,
cols = c("light grey", "mistyrose", "darkred"),
pt.size = 3,
add.alpha = TRUE,
ncol = 3, show.sb = FALSE,
pt.alpha = 0.1)

FeatureOverlay(se.cropped, features = "Dlx1",
sampleids = 2:4,
cols = c("light grey", "mistyrose", "darkred"),
pt.size = 3,
add.alpha = TRUE,
ncol = 3, show.sb = FALSE,
pt.alpha = 0.1)

FeatureOverlay(se.cropped, features = "Neurog2",
sampleids = 2:4,
cols = c("light grey", "mistyrose", "darkred"),
pt.size = 3,
add.alpha = TRUE,
ncol = 3, show.sb = FALSE,
pt.alpha = 0.1)

FeatureOverlay(se.cropped, features = "Eomes",
sampleids = 2:4,
cols = c("light grey", "mistyrose", "darkred"),
pt.size = 3,
add.alpha = TRUE,
ncol = 3, show.sb = FALSE,
pt.alpha = 0.1)

```

# G

>Fig6G: Violin plots illustrating selected genes from representative GO terms over-represented in GLU cells at P12.

```{r,fig.width=15,fig.height=4,fig.align="center"}
DefaultAssay(CLBNxs_Neurons_new_idents_V2)="RNA"
Idents(CLBNxs_Neurons_new_idents_V2)="GLUGABA_merge"

p1=VlnPlot(CLBNxs_Neurons_new_idents_V2, features = c("Id1"), idents=c("GLU_ident","GABA_ident"), cols=c("#133ea2","#b55b71"))+stat_compare_means(label = "p.signif", label.x = 1.5, ncol(1))+NoLegend()

p2=VlnPlot(CLBNxs_Neurons_new_idents_V2, features = c("Id2"), idents=c("GLU_ident","GABA_ident"), cols=c("#133ea2","#b55b71"))+stat_compare_means(label = "p.signif", label.x = 1.5, ncol(1))+NoLegend()

p3=VlnPlot(CLBNxs_Neurons_new_idents_V2, features = c("Zhx2"), idents=c("GLU_ident","GABA_ident"), cols=c("#133ea2","#b55b71"))+stat_compare_means(label = "p.signif", label.x = 1.5, ncol(1))+NoLegend()

p4=VlnPlot(CLBNxs_Neurons_new_idents_V2, features = c("Bmpr1a"), idents=c("GLU_ident","GABA_ident"), cols=c("#133ea2","#b55b71"))+stat_compare_means(label = "p.signif", label.x = 1.5, ncol(1))+NoLegend()

p5=VlnPlot(CLBNxs_Neurons_new_idents_V2, features = c("Ppp2r2b"), idents=c("GLU_ident","GABA_ident"), cols=c("#133ea2","#b55b71"))+stat_compare_means(label = "p.signif", label.x = 1.5, ncol(1))+NoLegend()

p6=VlnPlot(CLBNxs_Neurons_new_idents_V2, features = c("Btg2"), idents=c("GLU_ident","GABA_ident"), cols=c("#133ea2","#b55b71"))+stat_compare_means(label = "p.signif", label.x = 1.5, ncol(1))+NoLegend()

suppressWarnings(CombinePlots(plots = list(p1,p2,p3,p4,p5,p6),ncol=6)+NoLegend())
```

# H

>Fig6H: UMAP plot of GLU cells at P2, P12 and P22.

```{r,fig.width=5, fig.height=4,fig.align="center"}

V2_regression_lineage_GLU_without_GABA_and_oligos=readRDS("C:/Users/raine/OneDrive/Documents/BioInfo/lineage_GLU/V2_regression_lineage_GLU_without_GABA_and_oligos.rds")

Idents(V2_regression_lineage_GLU_without_GABA_and_oligos)="merge_clusters_to_use"

new_tp       <- c("P12", "P12", "P3", "P3", "P3", "P22", "P22")
names(new_tp)<- levels(V2_regression_lineage_GLU_without_GABA_and_oligos)
V2_regression_lineage_GLU_without_GABA_and_oligos         <- RenameIdents(V2_regression_lineage_GLU_without_GABA_and_oligos, new_tp)
V2_regression_lineage_GLU_without_GABA_and_oligos$new_tp   = V2_regression_lineage_GLU_without_GABA_and_oligos@active.ident
DimPlot(V2_regression_lineage_GLU_without_GABA_and_oligos,cols = c("#A94E66","#dcccd2","#510f3c"),pt.size=2)+NoLegend()

```

# I

>Fig6I: Feature plot indicating expression of Mik67 and Dcx.

```{r, fig.width=18, fig.height=4,fig.align="center"}

FeaturePlot(object = V2_regression_lineage_GLU_without_GABA_and_oligos,
            features = c("Dcx", "Mki67"),
            cols = c("#dedede", "#d1353e", "#3a8354"),
            blend = TRUE,
            combine=T, pt.size = 2, blend.threshold = 0.4, order=T)

```

# K

>Fig6K: Dotplot illustration enrichment of representative transcripts enriched in GLU cells at early (P2) or late (P12/P22) timepoints.

```{r, fig.width=7, fig.height=3.5,fig.align="center"}

new       <- c("P12_P22","P2", "P12_P22")
names(new)<- levels(V2_regression_lineage_GLU_without_GABA_and_oligos)
V2_regression_lineage_GLU_without_GABA_and_oligos         <- RenameIdents(V2_regression_lineage_GLU_without_GABA_and_oligos, new)
V2_regression_lineage_GLU_without_GABA_and_oligos$new   = V2_regression_lineage_GLU_without_GABA_and_oligos@active.ident

DefaultAssay(V2_regression_lineage_GLU_without_GABA_and_oligos)="RNA"
suppressWarnings(DotPlot(V2_regression_lineage_GLU_without_GABA_and_oligos, features = c("Tbr1","Neurod1","Neurod6","Neurod2","Dcx","Tubb3","Fbxw7","Jun","Notch1","Neurog2","Eomes","Cdkn1a","Mki67","Bmpr1a"),
        dot.scale = 10,cols = c("#572F4A","#DDD2D7"))+theme_grey()+RotatedAxis()+ xlab("")+ ylab("")+
  theme(axis.text.x = element_text(size=10, angle=45, hjust=1),
        axis.text.y = element_text(size=12,face="bold"),
        axis.title = element_text(size=14)))

```