---
title: "Script linked to Figure 5: Hopx and TGFβ/BMP signalling define pallial qNSCs quiescence/senescence."
subtitle: "Figure 5C, 5E, 5F, 5G"
output:
  rmdformats::downcute:
    self_contained: true
    default_style: "light"
    downcute_theme: "chaos"
---
***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load packages

```{r,warning=FALSE, include=T}
suppressWarnings(suppressMessages({
library(AUCell,quietly=T)
library(devtools,quietly=T)
library(tidyverse,quietly=T)
library(Seurat,quietly=T)
library(gridExtra,quietly=T)
library(gt,quietly=T)
library(glue,quietly=T)
library(scater,quietly=T)
library(tm,quietly=T)
library(gghalves,quietly=T)
library(broman,quietly=T)
library(viridis,quietly=T)
library(Cairo,quietly=T)
library(hrbrthemes,quietly=T)
library(ggplot2,quietly=T)
library(cowplot,quietly=T)
library(ggrepel,quietly=T)
library(plyr,quietly=T)
library(grid,quietly=T)
library(doRNG,quietly=T)
library(doSNOW,quietly=T)
library(mixtools,quietly=T)
library(SummarizedExperiment,quietly=T)
library(DT,quietly=T)
library(plotly,quietly=T)
library(NMF,quietly=T)
library(shiny,quietly=T)
library(rbokeh,quietly=T)
library(dynamicTreeCut,quietly=T)
library(R2HTML,quietly=T)
library(Rtsne,quietly=T)
library(zoo,quietly=T)
library(GSEABase,quietly=T)
library(magrittr, quietly=T)
library(imager, quietly=T)
library(EBImage, quietly=T)
library(STutility, quietly=T)
library(magrittr, quietly=T)
library(dplyr, quietly=T)
library(DT, quietly=T)
library(kableExtra, quietly=T)
library(ggpubr, quietly = T)
library(readr, quietly =T)
}))

```

# C

> Fig5C:  UMAP illustrating the pallial (red) and subpallial (blue) trajectories.

```{r,fig.width=6,fig.height=5,fig.align="center"}
same_umap_P2_P12_P22=readRDS("C:/Users/raine/OneDrive/Documents/BioInfo/newdataset/same_umap_P2_P12_P22.rds")
Idents(same_umap_P2_P12_P22)="orig.ident"
P12 = subset(same_umap_P2_P12_P22,idents = c("sdCLBNx1","sdCLBNx2"))

DimPlot(P12,group.by = "Full_detailed_clusters",cols = c("#ea878b","#DEDEDE","#99c3e1","#DEDEDE","#DEDEDE",
                                                         "#cef3fc","#DEDEDE","#DEDEDE","#f9bebe","#DEDEDE",
                                                         "#DEDEDE","#DEDEDE","#DEDEDE","#DEDEDE","#DEDEDE",
                                                         "#DEDEDE","#DEDEDE","#d26d6c","#4479b8","#d26d6c","#DEDEDE"),pt.size=1)+NoLegend()

```

# E

> Fig5E: Identification of key cellular states transitions along both trajectories.

```{r, warning=FALSE,fig.width=15, fig.height=12}

data_pallium <- read.csv("C:/Users/raine/OneDrive/Documents/BioInfo/papier figures et script/formated file for dotplot_Pallium.csv", sep =",", header = T, stringsAsFactors = FALSE)
data_pallium$group <- factor(data_pallium$group,levels = c("Pallial1", "Pallial2"))
data_pallium$name <- factor(data_pallium$name, levels = rev(unique(data_pallium$name)), ordered=TRUE)

Pallial <- ggplot(data_pallium, aes(x= group, y=name, size=NES, color=FDR, group=group))+ 
  scale_size(breaks = c(0.25, 0.5, 0.75, 1), range=c(0,8), limits = c(0.4,1)) + geom_point(alpha = 1,position = position_dodge(-7))+theme(panel.background = element_blank(),axis.text.y=element_text(size=15))
Pallial <- Pallial + scale_colour_gradient(low = "red", high = "lightgrey", space = "Lab")+xlab("")+ylab("")+ theme(plot.margin = margin(0.1,0.1,0.1,0.1, "cm"))


data_subpallium <- read.csv("C:/Users/raine/OneDrive/Documents/BioInfo/papier figures et script/formated file for dotplot_Subpallium.csv", sep =",", header = T, stringsAsFactors = FALSE)
data_subpallium$group <- factor(data_subpallium$group,levels = c("Subpallial1", "Subpallial2"))
data_subpallium$name <- factor(data_subpallium$name, levels = rev(unique(data_subpallium$name)), ordered=TRUE)

Subpallial <- ggplot(data_subpallium, aes(x= group, y=name, size=NES, color=FDR, group=group))+ 
  scale_size(breaks = c(0.25, 0.5, 0.75, 1), range=c(0,8), limits = c(0.4,1)) + geom_point(alpha = 1,position = position_dodge(16)) + theme(axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank(),panel.background = element_blank())
Subpallial <- Subpallial + scale_colour_gradient(low = "blue", high = "lightgrey", space = "Lab")+xlab("")+ylab("")+ theme(plot.margin = margin(0.1,0.1,0.1,0.1, "cm"))
```

```{r,fig.width=22,fig.height=8,warning = FALSE,fig.align="center"}

ggarrange(Pallial,Subpallial,legend = "top")+ theme(plot.margin = margin(0.1,0.1,0.1,0.1, "cm"))

```

# F

>Fig5F: Pseudotime analysis of Bmpr1a expression in the pallial and subpallial quiescence trajectories.

```{r, warning=FALSE,fig.width=5, fig.height=5,fig.align="center"}

pal <- readRDS("C:/Users/raine/Downloads/wetransfer_subpallial_no_alra_slingshot_qnscs_axis_subset-rds_2022-11-18_1148/pallial_no_alra_slingshot_qNSCs_axis_subset.rds")
subpal <- readRDS("C:/Users/raine/Downloads/wetransfer_subpallial_no_alra_slingshot_qnscs_axis_subset-rds_2022-11-18_1148/subpallial_no_alra_slingshot_qNSCs_axis_subset.rds")

subpal_counts = subpal@assays$RNA@data
subpal_counts = t(as.matrix(subpal_counts)) #switch row and columns
subpal_counts = as.data.frame(subpal_counts)

subpal_counts$pseudotime = subpal@meta.data$curve1
subpal_counts$pseudotime <- subpal_counts$pseudotime/max(subpal_counts$pseudotime)
subpal_counts$group <- "subpallial"


pal_counts = pal@assays$RNA@data
pal_counts = t(as.matrix(pal_counts)) #switch row and columns
pal_counts = as.data.frame(pal_counts)

pal_counts$pseudotime = pal@meta.data$curve1
pal_counts$pseudotime <- pal_counts$pseudotime/max(pal_counts$pseudotime)
pal_counts$group <- "pallial"

counts <- rbind(subpal_counts, pal_counts)
dataset_colors <- c(pallial = "#B31E2B", subpallial = "#2A4887")

genes <- c("Bmpr1a")

for (gene in genes) {
  p <- ggplot(counts, aes(pseudotime, .data[[gene]], group=group, col=group)) + 
    # geom_point() +
    geom_smooth(formula = y ~ x, size=2, method = "loess", se = FALSE) + 
    scale_color_manual(values=dataset_colors) + 
    labs(title = gene,
         color = "dataset")+NoLegend()+theme_classic()+coord_cartesian(ylim=c(0.50,1.50))
  plot(p)
  ggsave(paste0(gene,".png"), p, width = 5, height = 4)
}

```

# G

>Fig5G: Dot plot showing expression of TGFβ family type I receptor ligands transcripts per cell type.

```{r, warning=FALSE,fig.width=11, fig.height=5,fig.align="center"}
same_umap_P2_P12_P22=readRDS("C:/Users/raine/OneDrive/Documents/BioInfo/newdataset/same_umap_P2_P12_P22.rds")

Idents(same_umap_P2_P12_P22)="orig.ident"
P12_P2_test= subset(same_umap_P2_P12_P22,idents = c("p2","sdCLBNx1","sdCLBNx2"))
Idents(P12_P2_test)="predicted.simplifiedcelltype"

P12_P2_test@active.ident <- factor(P12_P2_test@active.ident, 
                            levels=c("P12_Ependymal cells", 
                                     "P12_qNSCs",
                                     "P12_aNSCs", 
                                     "P12_TAPs",
                                     "P12_Neuroblasts",
                                     "P12_Astrocytes", 
                                     "P12_Oligos", 
                                     "P12_Microglia", 
                                     "P12_Endothelial/Mural cells"))

suppressMessages(DotPlot(P12_P2_test, features = c("Bmp2", "Bmp4", "Gdf5", "Gdf6", "Bmp5", "Bmp6", "Bmp7", "Bmp8a", "Bmp8b", "Gdf15", "Amh", "Bmp15"), dot.scale = 10)+ 
  scale_colour_viridis() + RotatedAxis() + theme_grey() + theme(axis.title = element_text(size=15,colour = "white"),axis.text=element_text(size=12)))

```