---
title: "Script linked to Figure S4: Pallial and subpallial lineages coexist within the dorsal V-SVZ, related to Fig 2 & 3."
subtitle: "Figure S3A, S3B, S3G, S3H"
output:
  rmdformats::downcute:
    self_contained: true
    default_style: "light"
    downcute_theme: "chaos"
---
***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load packages

```{r,warning=FALSE, include=T}
suppressWarnings(suppressMessages({
library(tidyverse,quietly=T)
library(Seurat,quietly=T)
library(patchwork,quietly=T)
library(viridis,quietly=T)
library(scales,quietly=T)
library(treemap,quietly=T)
library(ggplot2,quietly=T)
library(ggpubr,quietly = T)
library(SCpubr,quietly=T)
}))
```

```{r,echo=FALSE}
P12_dSVZ_Seq=readRDS("P12_dSVZ_Seq/P12_dSVZ_Seq.rds")
```

# A

> FigS4A: Dual cells' average number of detected genes compared to pallial cells and subpallial cells.

```{r,fig.width=9,fig.height=5,fig.align="center"}
Idents(P12_dSVZ_Seq) = "Lineage"
###comment faire bar plot####
Idents(P12_dSVZ_Seq)="Simplified_clusters"
P12_dSVZ_Seq_subset=subset(P12_dSVZ_Seq,idents=c("qNSCs"))
Idents(P12_dSVZ_Seq_subset)="Lineage"
P12_dSVZ_Seq_subset=subset(P12_dSVZ_Seq_subset,idents=c("GLUCells","GABACells","DualCells"))
my_levels=c("DualCells","GLUCells","GABACells")
P12_dSVZ_Seq_subset$Lineage <- factor(x = P12_dSVZ_Seq_subset$Lineage, levels = my_levels)
Idents(P12_dSVZ_Seq_subset)="Lineage"
colors=c("GLUCells"="#ff0000","GABACells"="#0000ff","DualCells"="#fde725")
P1=SCpubr::do_BoxPlot(sample = P12_dSVZ_Seq_subset,feature = "nFeature_RNA",colors.use =colors,split.by="Lineage")+xlab("qNSCs")+NoLegend()

P12_dSVZ_Seq_subset=subset(P12_dSVZ_Seq,idents=c("aNSCs"))
Idents(P12_dSVZ_Seq_subset)="Lineage"
P12_dSVZ_Seq_subset=subset(P12_dSVZ_Seq_subset,idents=c("GLUCells","GABACells","DualCells"))
my_levels=c("DualCells","GLUCells","GABACells")
P12_dSVZ_Seq_subset$Lineage <- factor(x = P12_dSVZ_Seq_subset$Lineage, levels = my_levels)
Idents(P12_dSVZ_Seq_subset)="Lineage"
P2=SCpubr::do_BoxPlot(sample = P12_dSVZ_Seq_subset,feature = "nFeature_RNA",colors.use =colors,split.by="Lineage")+xlab("aNSCs")+NoLegend()

P12_dSVZ_Seq_subset=subset(P12_dSVZ_Seq,idents=c("TAPs"))
Idents(P12_dSVZ_Seq_subset)="Lineage"
P12_dSVZ_Seq_subset=subset(P12_dSVZ_Seq_subset,idents=c("GLUCells","GABACells","DualCells"))
my_levels=c("DualCells","GLUCells","GABACells")
P12_dSVZ_Seq_subset$Lineage <- factor(x = P12_dSVZ_Seq_subset$Lineage, levels = my_levels)
Idents(P12_dSVZ_Seq_subset)="Lineage"
P3=SCpubr::do_BoxPlot(sample = P12_dSVZ_Seq_subset,feature = "nFeature_RNA",colors.use =colors,split.by="Lineage")+xlab("TAPs")+NoLegend()

P1+P2+P3
```

# B

> FigS4B: Dual cells proportions within V-SVZ cell types.

```{r,message=FALSE,warning=FALSE}
plot.data <- P12_dSVZ_Seq@meta.data %>%
  dplyr:::select(Lineage, cluster   = Simplified_clusters) %>%
  group_by(cluster, Lineage ) %>%
  summarise(count      = n()) %>%
  mutate(clust_total   = sum(count)) %>%
  mutate(clust_prop    = count / clust_total) %>%
  group_by(Lineage) %>%
  mutate(dataset_total = sum(count)) %>%
  ungroup() %>%
  mutate(dataset_prop  = count / dataset_total)

plt.data= plot.data %>% subset(cluster %in% c("qNSCs","aNSCs","TAPs") & Lineage=="DualCells") %>% mutate(cluster = fct_reorder(cluster, -desc(clust_prop)))%>% mutate(across((clust_prop),function(x) x*100))

#Distribution % in each clusters
ggplot(plt.data, aes(x = cluster, y = clust_prop, fill=cluster)) +
  geom_col()+scale_fill_manual(values=c("#c29200" ,"#9aa500","#45c2b2"))+theme_pubr()+NoLegend()
```

# G 

> FigS2G: UMAP plots of integrated datasets presented in Fig 3A, with identity of Loo datasets.

```{r}
Loo_Marcy=readRDS("Loo_Marcy/Loo_Marcy.rds")
Idents(Loo_Marcy) = "Simplified_loo_clusters"
DimPlot(Loo_Marcy,cols=viridis(19),na.value = "#DEDEDE",group.by = "Simplified_loo_clusters")
```

# H 

> FigS2H: UMAP plots of integrated datasets presented in Fig 3A, with pallial and subpallial cells from current P12 dataset.

```{r}
Idents(Loo_Marcy) = "Lineages"
colors=c("GLUCells"="#ff0000","GABACells"="#0000ff","DualCells"="#fde725","NACells"="#DEDEDE")
DimPlot(Loo_Marcy,cols=colors,group.by = "Lineages",na.value="grey")

```
