---
title: "Script linked to Fig S5. Pallial and subpallial qNSCs differ in their transcriptional profile and transcript content, related to Figure 4"
subtitle: "Figure 5A, 5B, 5C, 5D, 5E, 5F, 5G, 5H"
output:
  rmdformats::downcute:
    self_contained: true
    default_style: "light"
    downcute_theme: "chaos"
---
***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load packages

```{r,warning=FALSE, include=T}
suppressWarnings(suppressMessages({
library(AUCell,quietly=T)
library(devtools,quietly=T)
library(tidyverse,quietly=T)
library(Seurat,quietly=T)
library(gridExtra,quietly=T)
library(gt,quietly=T)
library(glue,quietly=T)
library(scater,quietly=T)
library(tm,quietly=T)
library(gghalves,quietly=T)
library(broman,quietly=T)
library(viridis,quietly=T)
library(Cairo,quietly=T)
library(hrbrthemes,quietly=T)
library(ggplot2,quietly=T)
library(cowplot,quietly=T)
library(ggrepel,quietly=T)
library(plyr,quietly=T)
library(grid,quietly=T)
library(doRNG,quietly=T)
library(doSNOW,quietly=T)
library(mixtools,quietly=T)
library(SummarizedExperiment,quietly=T)
library(DT,quietly=T)
library(plotly,quietly=T)
library(NMF,quietly=T)
library(shiny,quietly=T)
library(rbokeh,quietly=T)
library(dynamicTreeCut,quietly=T)
library(R2HTML,quietly=T)
library(Rtsne,quietly=T)
library(zoo,quietly=T)
library(GSEABase,quietly=T)
library(magrittr, quietly=T)
library(imager, quietly=T)
library(EBImage, quietly=T)
library(STutility, quietly=T)
library(magrittr, quietly=T)
library(dplyr, quietly=T)
library(DT, quietly=T)
library(kableExtra, quietly=T)
library(ggpubr, quietly = T)
}))
```


```{r, warning=FALSE}
P12_dSVZ_Seq=readRDS("P12_dSVZ_Seq/P12_dSVZ_Seq.rds")
```

# A

> FigS5A: UMAP highlighting 2 qNSCs subclusters.

```{r,fig.width=5,fig.height=4,fig.align="center"}
Idents(P12_dSVZ_Seq) = "Full_clusters"
DimPlot(P12_dSVZ_Seq, label = F,label.size = 4, pt.size = 1,repel=T,cols= c("#DEDEDE","#DEDEDE","#6F4D0F","#DEDEDE","#C59226","#DEDEDE","#DEDEDE","#DEDEDE","#DEDEDE","#DEDEDE","#DEDEDE","#DEDEDE","#DEDEDE","#DEDEDE","#DEDEDE","#DEDEDE"))+NoLegend()

```

# B

> FigS5B: Violin plots illustrating expression of transcription factors/transcriptional regulators enriched in qNSC1 and qNSC2

```{r,fig.width=30,fig.height=15,fig.align="center"}
DefaultAssay(P12_dSVZ_Seq)="RNA"
Idents(P12_dSVZ_Seq) = "Full_clusters"


p1=VlnPlot(P12_dSVZ_Seq, features = c("Emx1"), idents=c("qNSC1","qNSC2"), cols=c("#6F4D0F","#C59226"))+stat_compare_means(label = "p.signif", label.x = 1.5)+NoLegend()

p2=VlnPlot(P12_dSVZ_Seq, features = c("Zic5"), idents=c("qNSC1","qNSC2"), cols=c("#6F4D0F","#C59226"))+stat_compare_means(label = "p.signif", label.x = 1.5)+NoLegend()

p3=VlnPlot(P12_dSVZ_Seq, features = c("Tfap2c"), idents=c("qNSC1","qNSC2"), cols=c("#6F4D0F","#C59226"))+stat_compare_means(label = "p.signif", label.x = 1.5)+NoLegend()

p4=VlnPlot(P12_dSVZ_Seq, features = c("Dmrta2"), idents=c("qNSC1","qNSC2"), cols=c("#6F4D0F","#C59226"))+stat_compare_means(label = "p.signif", label.x = 1.5)+NoLegend()

p5=VlnPlot(P12_dSVZ_Seq, features = c("Fezf2"), idents=c("qNSC1","qNSC2"), cols=c("#6F4D0F","#C59226"))+stat_compare_means(label = "p.signif", label.x = 1.5)+NoLegend()

p6=VlnPlot(P12_dSVZ_Seq, features = c("Hopx"), idents=c("qNSC1","qNSC2"), cols=c("#6F4D0F","#C59226"))+stat_compare_means(label = "p.signif", label.x = 1.5)+NoLegend()

p7=VlnPlot(P12_dSVZ_Seq, features = c("Id1"), idents=c("qNSC1","qNSC2"), cols=c("#6F4D0F","#C59226"))+stat_compare_means(label = "p.signif", label.x = 1.5)+NoLegend()

p8=VlnPlot(P12_dSVZ_Seq, features = c("Id2"), idents=c("qNSC1","qNSC2"), cols=c("#6F4D0F","#C59226"))+stat_compare_means(label = "p.signif", label.x = 1.5)+NoLegend()

p9=VlnPlot(P12_dSVZ_Seq, features = c("Id3"), idents=c("qNSC1","qNSC2"), cols=c("#6F4D0F","#C59226"))+stat_compare_means(label = "p.signif", label.x = 1.5)+NoLegend()

p10=VlnPlot(P12_dSVZ_Seq, features = c("Id4"), idents=c("qNSC1","qNSC2"), cols=c("#6F4D0F","#C59226"))+stat_compare_means(label = "p.signif", label.x = 1.5)+NoLegend()

p11=VlnPlot(P12_dSVZ_Seq, features = c("Gsx2"), idents=c("qNSC1","qNSC2"), cols=c("#6F4D0F","#C59226"))+stat_compare_means(label = "p.signif", label.x = 1.5)+NoLegend()

p12=VlnPlot(P12_dSVZ_Seq, features = c("Six3"), idents=c("qNSC1","qNSC2"), cols=c("#6F4D0F","#C59226"))+stat_compare_means(label = "p.signif", label.x = 1.5)+NoLegend()

p13=VlnPlot(P12_dSVZ_Seq, features = c("Rorb"), idents=c("qNSC1","qNSC2"), cols=c("#6F4D0F","#C59226"))+stat_compare_means(label = "p.signif", label.x = 1.5)+NoLegend()

p14=VlnPlot(P12_dSVZ_Seq, features = c("Eno1"), idents=c("qNSC1","qNSC2"), cols=c("#6F4D0F","#C59226"))+stat_compare_means(label = "p.signif", label.x = 1.5)+NoLegend()

suppressWarnings(CombinePlots(plots = list(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,p11,p12,p13,p14),ncol=5))
```

# C

> FigS5C: Violin plots illustrating higher transcript content of qNSC1 when compared qNSC2.

```{r,fig.width=10,fig.height=6,fig.align="center"}
DefaultAssay(P12_dSVZ_Seq)="RNA"
Idents(P12_dSVZ_Seq) = "Full_clusters"

VlnPlot(P12_dSVZ_Seq, features = c("nFeature_RNA"), idents=c("qNSC1","qNSC2"), cols=c("#6F4D0F","#C59226"), pt.size=1.5)+stat_compare_means(label = "p.signif", label.x = 1.5)
```

# D

> FigS5D: UMAP plot showing enrichment of HopxHight cells in qNSC1, while HopxLow cells are mainly associated to qNSC2

```{r,fig.width=6,fig.height=4,fig.align="center"}

qNSCs <- subset(P12_dSVZ_Seq, idents=c("qNSC1","qNSC2"))
DefaultAssay(qNSCs)="RNA"
HopxHigh=WhichCells(qNSCs,expression=Hopx>2.5)
HopxLow=WhichCells(qNSCs,expression=Hopx<0.4)

qNSCs <- SetIdent(qNSCs, cells =HopxHigh , value = "HopxHigh") #first transform the cells into an ident
qNSCs <- SetIdent(qNSCs, cells =HopxLow , value = "HopxLow") #first transform the cells into an ident
qNSCs[["new_idents_Hopx"]] = Idents(qNSCs) #then write a new column called new_clusters
Idents(qNSCs) = "new_idents_Hopx"
DimPlot(qNSCs, cols=c("#0000ff","#ff0000","#6f4f00","#c99800"), pt.size=1)
```

```{r}
DefaultAssay(qNSCs)="RNA"
VlnPlot(qNSCs, features = "nFeature_RNA", idents=c("HopxHigh","HopxLow") ,cols=c("#0000ff","#ff0000"))+stat_compare_means(label = "p.signif", label.x = 1.5)
```

# F&G

> FigS5F to G: Violin plots illustrating the enrichment of genes associated to deep quiescence 

```{r,fig.width=10,fig.height=4,fig.align="center"}
DefaultAssay(P12_dSVZ_Seq)="RNA"
Idents(P12_dSVZ_Seq) = "Full_clusters"

p1=VlnPlot(P12_dSVZ_Seq, features = c("Cdkn1b"), idents=c("qNSC1","qNSC2"), cols=c("#6F4D0F","#C59226"))+stat_compare_means(label = "p.signif", label.x = 1.5)+NoLegend()

p2=VlnPlot(P12_dSVZ_Seq, features = c("Bmpr1a"), idents=c("qNSC1","qNSC2"), cols=c("#6F4D0F","#C59226"))+stat_compare_means(label = "p.signif", label.x = 1.5)+NoLegend()

p3=VlnPlot(P12_dSVZ_Seq, features = c("Egfr"), idents=c("qNSC1","qNSC2"), cols=c("#6F4D0F","#C59226"))+stat_compare_means(label = "p.signif", label.x = 1.5)+NoLegend()

p4=VlnPlot(P12_dSVZ_Seq, features = c("Cd9"), idents=c("qNSC1","qNSC2"), cols=c("#6F4D0F","#C59226"))+stat_compare_means(label = "p.signif", label.x = 1.5)+NoLegend()

suppressWarnings(CombinePlots(plots = list(p1,p2,p3,p4),ncol=4))
```

# H

> FigS5H : calculation of a “deep quiescence” AUC score 

```{r,warning=FALSE,include=F}
Deep_Quiescence = c("Notch2",
        "Foxg1",
        "Prom1",
        "Id1",
        "Foxo3",
        "Sox2",
        "Sox9",
        "Pdgfra",
        "Nr2e1",
        "Ascl1", "Nfix")

data_matrix <- GetAssayData(P12_dSVZ_Seq, assay = "RNA")
data_matrix <- as.matrix(data_matrix)

deepquiesencesets = Deep_Quiescence
deepquiesencesets <- GeneSet(deepquiesencesets, setName= "deepquiesencesets")
CellTypes10 <- GeneSetCollection(deepquiesencesets)
ranking <- AUCell_buildRankings(data_matrix)

AUC_10 <- AUCell_calcAUC(CellTypes10, ranking)
rankings10 <- apply(AUC_10@assays@data@listData[["AUC"]], 1, function(x) (x-min(x))/(max(x)-min(x))) 
rankings10 <- as.data.frame(rankings10)
rankings10$class <- as.data.frame(P12_dSVZ_Seq@meta.data$Full_clusters, row.names = Cells(P12_dSVZ_Seq))[row.names(rankings10),]
```

```{r,warning=FALSE,fig.align="center"}

my_comparisons <- list(c("qNSC1", "qNSC2"))
deepquiescent <- c("qNSC1" = "#75540D" , "qNSC2"="#CB9B20")
suppressMessages(suppressWarnings(rankings10 %>%
  filter(class == "qNSC1" |
          class == "qNSC2") %>%
  ggplot(aes(x = class, y = deepquiesencesets,fill=class)) +
  geom_violin()+ggtitle("AUCell score deep quiescence markers")+stat_compare_means(label="p.signif",comparisons = my_comparisons)+ ylab("AUCell score")+theme_ipsum()+ scale_x_discrete("class", limits = c("qNSC1", "qNSC2"), labels=c("qNSC1","qNSC2")) + scale_fill_manual(values=deepquiescent)+NoLegend()))+ geom_jitter()
```